---
title: "Palmer Penguins"
subtitle: "[TidyTuesday](https://github.com/rfordatascience/tidytuesday)"
author: "Akil Elkamel"
date: "7/28/2020"
output:
  html_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Data

The data this week comes from Dr. Kristen Gorman by way of the palmerpenguins R package by Dr. Kristen Gorman, Dr. Allison Horst, and Dr. Alison Hill.

Their palmerpenguins packagedown site and corresponding GitHub Repo has all the details, which I will duplicate some of below.

You can install their package for reproducible use via install.packages("palmerpenguins").

The goal of palmerpenguins is to provide a great dataset for data exploration & visualization, as an alternative to iris.

We gratefully acknowledge Palmer Station LTER and the US LTER Network. Special thanks to Marty Downs (Director, LTER Network Office) for help regarding the data license & use.

They've bundled both the raw data and the cleaned data together, which I have also included here.

The main measurements are body mass, culmen (bill) length, bill depth, and flipper length.

Get the data here

```{r eval=FALSE}
# Get the Data

# Read in with tidytuesdayR package 
# Install from CRAN via: install.packages("tidytuesdayR")
# This loads the readme and all the datasets for the week of interest

# Either ISO-8601 date or year/week works!

tuesdata <- tidytuesdayR::tt_load('2020-07-28')
tuesdata <- tidytuesdayR::tt_load(2020, week = 31)

penguins <- tuesdata$penguins

# Or read in the data manually

penguins.csv <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins.csv')

penguins_raw.csv <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins_raw.csv')
```

```{r include=FALSE, eval=FALSE}
tuesdata <- tidytuesdayR::tt_load('2020-07-28')
penguins <- tuesdata$penguins 
```

```{r}
library(palmerpenguins)

```


## Data Dictionary

**penguins.csv**


| variable          | class   | description                                              |
| :---------------- | :------ | :------------------------------------------------------- |
| species           | integer | Penguin species (Adelie, Gentoo, Chinstrap)              |
| island            | integer | Island where recorded (Biscoe, Dream, Torgersen)         |
| bill_length_mm    | double  | Bill length in millimeters (also known as culmen length) |
| bill_depth_mm     | double  | Bill depth in millimeters (also known as culmen depth)   |
| flipper_length_mm | integer | Flipper length in mm                                     |
| body_mass_g       | integer | Body mass in grams                                       |
| sex               | integer | sex of the animal                                        |
| year              | integer | year recorded                                            |


**penguins_raw.csv**

| variable            | class     | description                                           |
| :------------------ | :-------- | :---------------------------------------------------- |
| studyName           | character | Study name                                            |
| Sample Number       | double    | Sample id                                             |
| Species             | character | Species of penguin                                    |
| Region              | character | Region where recorded                                 |
| Island              | character | Island where recorded                                 |
| Stage               | character | Stage of egg                                          |
| Individual ID       | character | Individual penguin ID                                 |
| Clutch Completion   | character | Egg clutch completion                                 |
| Date Egg            | double    | Date of egg                                           |
| Culmen Length (mm)  | double    | culmen length in mm (beak length)                     |
| Culmen Depth (mm)   | double    | culmen depth in mm (beak depth)                       |
| Flipper Length (mm) | double    | Flipper length in mm                                  |
| Body Mass (g)       | double    | Body mass in g                                        |
| Sex                 | character | Sex of the penguin                                    |
| Delta 15 N (o/oo)   | double    | Blood isotopic Nitrogen - used for dietary comparison |
| Delta 13 C (o/oo)   | double    | Blood isotopic Carbon - used for dietary comparison   |
| Comments            | character | Miscellaneous comments                                |


## Cleaning Script

No cleaning script today, feel free to work with the pre-cleaned data or try your hand at the raw data!



# Exploratory Data Analysis

```{r include=FALSE}
library(tidyverse)
```


## Dataset discovery

A first look to our data:

The dataset is composed by 344 observations (penguins) described by eight variables each. The varibales are presented previously in the table *penguins.csv*

```{r}
penguins %>% 
    head()
```

A summary of the data:

```{r}
penguins %>% 
    summary()
```

We have three species of penguins distriuted as follow:

```{r}
penguins %>% 
    count(species, sort = TRUE)
```

Our penguins live in three islands

```{r}
penguins %>% 
    count(island, sort = TRUE)
```

We can see here that *Chinstrap* penguins live only in the *Dream* island, and the same thing for the *Gentoo* penguins which live only in the *Biscoe* island, while the *Adelie* penguins live in the three islands. We notice also that the *Torgerson* island contain only *Adelie* penguins.


```{r}
penguins %>% 
    count(species, island)
```


```{r}
penguins %>%
	ggplot(aes(island, fill = species)) +
	geom_bar()
```


## Data taming

For the species, island and sex columns, we see that their types are `character`, it must be a `factor`.

```{r}
penguins <- penguins %>% 
    mutate_if(is.character, factor)

penguins %>% 
    summary()
```

## Missing data

We see that we have 11 observations without sex value, among of them two observations are missing also *bill_length_mm*, *bill_depth_mm*, *flipper_length_mm*, and *body_mass_g*.

```{r}
penguins %>% 
    is.na() %>% 
    colSums()
```


```{r}
# Load the VIM package
library(VIM)

# Draw an aggregation plot of biopics
penguins %>% 
    aggr(combined = TRUE, numbers = TRUE)
```


```{r}
library(visdat)
penguins %>% 
    vis_miss()
```


```{r}
penguins %>% 
    filter(is.na(sex))
```



In the following we will use penguins without missing data.

```{r}
penguins <- penguins%>% 
    filter(!is.na(sex))
```

## Data imputation

```{r eval=FALSE}
library(missForest)

penguins <- penguins %>% 
    mutate(sex_na = is.na(sex))

penguins <- as.data.frame(penguins)

imp_res <- missForest(penguins)

imp_err <- imp_res$OOBerror
print(imp_err)

# Pair-wise technique

imp_res <- missForest(penguins, variablewise = TRUE)

per_variable_errors <- imp_res$OOBerror
print(per_variable_errors)

names(per_variable_errors) <- paste(names(penguins),
                                    names(per_variable_errors), 
                                    sep = "_")

# Print the renamed errors
print(per_variable_errors)

imp_res
```

# Visual exploration

A first visual exploration that shows the distribution of penguins according to "*bill_depth_mm*" vs "*bill_length_mm*".

```{r}
penguins %>% 
    filter(!is.na(sex)) %>% 
    ggplot(aes(x = bill_length_mm, y = bill_depth_mm, col = species, size = body_mass_g)) +
        geom_point() + 
        facet_wrap(~ sex)
```


A histogram plot of the different measures that characterize penguins.

```{r}
penguins %>% 
    pivot_longer(bill_length_mm:body_mass_g,
                 names_to = "measure",
                 values_to = "value") %>% 
    ggplot(aes(x = value, fill = species)) +
    geom_histogram(bins = 15) +
    facet_wrap(~ measure, scales = "free_x")
```

And here is the distribution of the different measures:

```{r}
penguins %>% 
    pivot_longer(bill_length_mm:body_mass_g,
                 names_to = "measure",
                 values_to = "value") %>% 
    ggplot(aes(x = value, fill = species)) +
    geom_density(alpha = 0.4) +
    facet_wrap(~ measure, scales = "free")
```


As we see in these boxplots, the *Gento* penguins can be distinguished easily using "*bill_depth_mm*", "*flipper_length_mm*" and "*body_mass_g*".

The "*bill_length_mm*"

```{r}
penguins %>% 
    pivot_longer(bill_length_mm:body_mass_g,
                 names_to = "measure",
                 values_to = "value") %>% 
    ggplot(aes(x = species, y = value)) +
    geom_boxplot() +
    facet_wrap(~ measure, scales = "free_y")
```




# Modeling

We will use in this step `tidymodels`.

```{r message=FALSE}
library(tidymodels)
```

We minimize our dataset to get only necessary columns:

```{r}
penguins <- penguins %>% 
  select(species, bill_length_mm:body_mass_g)
```


## Train / Test data split

```{r}
train_test_split <- penguins %>% 
  initial_split(prop = 4/5, strata = species)

penguins_train <- training(train_test_split)
penguins_test <- testing(train_test_split)
```

## Evaluation metrics

```{r}
ev_metrics <- metric_set(accuracy, kap, roc_auc)
```

```{r}
models_res <- testing(train_test_split) %>% 
  select(actual = species)
```


## Multinomial Regression Model


```{r}
multinom_reg_clf <- multinom_reg(mode = "classification", penalty = 0) %>% 
  set_engine("glmnet") %>% 
  fit(formula = species ~ bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g,
      data = training(train_test_split))

pred <- testing(train_test_split) %>% 
  predict(multinom_reg_clf, new_data = .) 

models_res <- models_res %>% 
  bind_cols(pred) %>% 
  rename(multinom_reg = .pred_class)
```


## K Nearest Neighbors

```{r}
knn_clf <- nearest_neighbor(mode = "classification",
                            neighbors = 13) %>% 
  set_engine("kknn") %>% 
  fit(formula = species ~ bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g,
      data = training(train_test_split))

pred <- testing(train_test_split) %>% 
  predict(knn_clf, new_data = .) 

models_res <- models_res %>% 
  bind_cols(pred) %>% 
  rename(knn = .pred_class)
```

## Support Vector Machine

```{r}
svm_clf <- svm_poly(mode = "classification") %>% 
  set_engine("kernlab") %>% 
  fit(formula = species ~ bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g,
      data = training(train_test_split))

pred <- testing(train_test_split) %>% 
  predict(svm_clf, new_data = .) 

models_res <- models_res %>% 
  bind_cols(pred) %>% 
  rename(svm = .pred_class)

```


## Decision Tree

```{r}
dt_clf <- decision_tree(mode = "classification") %>% 
  set_engine("rpart") %>% 
  fit(formula = species ~ bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g,
      data = training(train_test_split))

pred <- testing(train_test_split) %>% 
  predict(dt_clf, new_data = .) 

models_res <- models_res %>% 
  bind_cols(pred) %>% 
  rename(dt = .pred_class)

```

## Random Forest

```{r}
rf_clf <- rand_forest(mode = "classification") %>% 
  set_engine("ranger") %>% 
  fit(formula = species ~ bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g,
      data = training(train_test_split))

pred <- testing(train_test_split) %>% 
  predict(rf_clf, new_data = .) 

models_res <- models_res %>% 
  bind_cols(pred) %>% 
  rename(rf = .pred_class)

```



```{r}
models_res
```





```{r}
table(models_res$actual, models_res$multinom_reg)
table(models_res$actual, models_res$knn)
table(models_res$actual, models_res$svm)
table(models_res$actual, models_res$dt)
table(models_res$actual, models_res$rf)
```



## Advanced


### Bootstrap

```{r}
set.seed(1234)
penguins_boot <- bootstraps(penguins_train, strata = species)
```


### Workflow

```{r}
penguins_wf <- workflow() %>% 
  add_formula(species ~ .)

penguins_wf
```

### Add models to a workflow


We define here some models specifications:

 * KNN 
 * Random Forest

```{r}
knn_spec <- nearest_neighbor() %>% 
  set_mode("classification") %>% 
  set_engine("kknn")

rf_spec <- rand_forest() %>% 
  set_mode("classification") %>% 
  set_engine("ranger")
```
 
```{r}
knn_rs <- penguins_wf %>% 
  add_model(knn_spec) %>% 
  fit_resamples(
    resamples = penguins_boot,
    control = control_resamples(
      save_pred = TRUE,
      verbose = TRUE
    )
  )
```


```{r}
rf_rs <- penguins_wf %>% add_model(rf_spec) %>% 
  fit_resamples(
    resamples = penguins_boot,
    control = control_resamples(
      save_pred = TRUE,
      verbose = TRUE
    )
  )
```

### Evaluate models

```{r}
collect_metrics(knn_rs)
collect_metrics(rf_rs)
```


```{r}
knn_rs %>% 
  conf_mat_resampled()
```


### ROC-AUC

With autoplot

```{r}
knn_rs %>% 
  collect_predictions() %>%
  group_by(id) %>% 
  roc_curve(species, .pred_Adelie:.pred_Gentoo) %>% 
  autoplot()
```


Manually with ggplot

```{r}
knn_rs %>% 
  collect_predictions() %>%
  group_by(id) %>% 
  roc_curve(species, .pred_Adelie:.pred_Gentoo) %>% 
  ggplot(aes(1-specificity, sensitivity, color = id)) + 
  geom_abline(lty = 2, color = "gray80", size = 1.25) +
  geom_path(show.legend = FALSE, alpha = 0.7, size = 1.2) +
  coord_equal() +
  facet_wrap(~ .level)
```


## Cross-Validation

```{r}
penguins_train_cv <- penguins_train %>% 
  rsample::vfold_cv(strata = species)

rf <- rand_forest() %>% 
  set_mode("classification") %>% 
  set_engine("randomForest")
```

```{r}
wf <- workflow() %>% 
  add_formula(species ~ .)

res <- wf %>% 
  add_model(rf) %>% 
  fit_resamples(
    resamples = penguins_train_cv,
    metrics = metric_set(accuracy, kap, roc_auc)
  )

res %>% 
  collect_metrics()
```


```{r}
res %>% 
  collect_metrics() %>% 
  ggplot(aes(mean, .metric)) + 
  geom_point() +
  geom_errorbarh(aes(xmin = mean - std_err,
                     xmax = mean + std_err))
```


