---
title: "First tidymodels Example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Import packages

```{r}
library(tidyverse)
library(tidymodels)
```


## Import data

```{r}
penguins_raw <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins.csv')
```


```{r}
penguins <- penguins_raw %>% 
  filter(!is.na(sex))
```


## Split data into training and test datasets

```{r}
set.seed(4254)
split <- initial_split(penguins_raw, prop = 0.8, strata = species)

peng_training <- training(split)
peng_test <- testing(split)

peng_training %>% 
  is.na() %>% 
  colSums()
peng_test %>% 
  is.na() %>% 
  colSums()

peng_kfolds <- vfold_cv(peng_training)
```

## Prepare the recipe


```{r}
peng_rec <- recipe(species ~ ., data = penguins) %>% 
  step_knnimpute(sex) %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_normalize(all_numeric()) # to remove

peng_rec %>% prep() %>% juice()
```


## Create the models


By using the parsnip package to define ML models. Parsnip offers a general interface for most the ML algorithms.

```{r}

knn_spec <- nearest_neighbor(
  neighbors = tune()
) %>% set_mode("classification") %>% 
  set_engine("kknn")

knn_spec

rf_spec <- rand_forest(
  mtry = tune(),
  trees = tune(),
  min_n = tune()
) %>% set_mode("classification") %>% 
  set_engine("ranger")

rf_spec
```


## Fine tuning the models


Using the tune and dials packages.

 * dials: create the params grid
 * tune: will tune the params according to the combinations of the grid
 

Here we see the list of parameters for our models

```{r}
parameters(knn_spec)
parameters(rf_spec)
```

Here we use `grid_regular()` from dials packages and `parameters()` to see the list of grid parameters.

```{r}
grid_regular(parameters(knn_spec), levels = 10)
#grid_regular(parameters(rf_spec), levels = 10)
```



```{r}
peng_wf <- workflow() %>%
  add_recipe(peng_rec) %>%
  add_model(knn_spec)
```


```{r}
library(doParallel)
doParallel::registerDoParallel()

set.seed(345)
tune_res <- tune_grid(
  peng_wf,
  resamples = peng_kfolds,
  grid = 20
)

tune_res
```



