---
title: "First tidymodels Example"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Import packages

```{r}
library(tidyverse)
library(tidymodels)
```


## Import data

```{r}
penguins_raw <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins.csv')
```


```{r}
penguins_raw %>% 
  is.na() %>% 
  colSums()

penguins_raw %>% 
  head(10)

penguins <- penguins_raw %>% 
  select(-c(sex, year)) %>% 
  na.omit()

penguins %>% 
  head(10)
```


```{r}
library(skimr)
skim(penguins)
```

## Split data into training and test datasets

```{r}
set.seed(4254)
split <- initial_split(penguins, prop = 0.8, strata = species)

peng_training <- training(split)
peng_test <- testing(split)

peng_training %>% 
  count(species, sort = TRUE)


peng_test %>% 
  count(species, sort = TRUE)

peng_kfolds <- vfold_cv(peng_training)
```

## Prepare the recipe


```{r}
peng_rec <- recipe(species ~ ., data = penguins) %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_normalize(all_numeric()) %>% # to remove
  prep(training = peng_training)

peng_rec  %>% juice()
```


## Create the models


By using the parsnip package to define ML models. Parsnip offers a general interface for most the ML algorithms.

```{r}
knn_spec <- nearest_neighbor(
  neighbors = tune()
) %>% set_mode("classification") %>% 
  set_engine("kknn")

knn_spec

rf_spec <- rand_forest(
  mtry = 5,
  trees = tune(),
  min_n = tune()
) %>% set_mode("classification") %>% 
  set_engine("ranger")

rf_spec
```


## Fine tuning the models


Using the tune and dials packages.

 * dials: create the params grid
 * tune: will tune the params according to the combinations of the grid
 

Here we see the list of parameters for our models

```{r}
parameters(knn_spec)
parameters(rf_spec)
```

Here we use `grid_regular()` from dials packages and `parameters()` to see the list of grid parameters.

```{r}
grid_regular(parameters(knn_spec), levels = 10)
grid_regular(parameters(rf_spec), levels = 10)
```



```{r}
peng_wf <- workflow() %>%
  add_recipe(peng_rec) %>%
  add_model(rf_spec)
```


```{r}
library(doParallel)
doParallel::registerDoParallel()

set.seed(345)
tune_res <- tune_grid(
  peng_wf,
  resamples = peng_kfolds,
  grid = 20
)

tune_res
```

```{r}
tune_res %>%
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  select(mean, trees, min_n) %>%
  pivot_longer(trees:min_n,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "AUC")
```



```{r}
best_auc <- select_best(tune_res, "roc_auc")

final_rf <- finalize_model(
  rf_spec,
  best_auc
)

final_rf
```





## KNN tuning and choosing the best model

```{r}
peng_wf <- workflow() %>%
  add_recipe(peng_rec) %>%
  add_model(knn_spec)
```


```{r}
library(doParallel)
doParallel::registerDoParallel()

set.seed(345)
tune_res <- tune_grid(
  peng_wf,
  resamples = peng_kfolds,
  grid = 20
)

tune_res
```

```{r}
tune_res %>%
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  select(mean, neighbors) %>%
  ggplot(aes(neighbors, mean)) +
  geom_point(show.legend = FALSE) +
  labs(x = NULL, y = "AUC")
```



```{r}
best_auc <- select_best(tune_res, "roc_auc")

final_knn <- finalize_model(
  knn_spec,
  best_auc
)

final_knn
```

## Variable importance

```{r}
rf_prep <- prep(peng_rec)
juiced <- juice(rf_prep)
```


```{r}
library(vip)

final_rf %>%
  set_engine("ranger", importance = "permutation") %>%
  fit(species ~ .,
    data = juice(rf_prep)
  ) %>%
  vip(geom = "point")
```




```{r}
final_wf <- workflow() %>%
  add_recipe(peng_rec) %>%
  add_model(final_rf)

final_res <- final_wf %>%
  last_fit(split)

final_res %>%
  collect_metrics()
```



```{r}
final_res %>%
  collect_predictions() %>%
  mutate(correct = case_when(
    species == .pred_class ~ "Correct",
    TRUE ~ "Incorrect"
  )) %>%
  bind_cols(peng_test) %>%
  ggplot(aes(bill_length_mm, bill_depth_mm, color = correct)) +
  geom_point(size = 2, alpha = 1) +
  labs(color = NULL) +
  scale_color_manual(values = c("gray80", "darkred"))
```















